%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Quick start with JACKS}
---
title: "QuickStart JACKS"
author: "Felicity Allen, Leopold Parts"
date: "2/14/2018"
output: 
    html_document:
      toc: true
      theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jacks)
```
### Load Input
The expected input is a file containing all the raw gRNA counts and another file containing the replicate to sample specifications, please see 
https://raw.githubusercontent.com/felicityallen/JACKS/master/jacks/README.txt
for a full explanation of the different files and fields (which are common with the python version of JACKS).
The gene specification can optionally be an additional file, or the count file can be used for this if it has a column specifying the gene (this is the default).
The following functions load these counts, and computes the normalized mean and standard deviation of each gRNA across all available repliates for each sample.
```{r counts, warning=FALSE, message=FALSE}
    count_file = system.file("extdata", "example_count_data.tab", package="jacks", mustWork=TRUE)
    sample_spec_file = system.file("extdata", "example_repmap.tab", package="jacks", mustWork=TRUE)
    lfc = read_counts_from_spec_files(count_file, sample_spec_file, replicate_col="Replicate", sample_col="Sample", gene_spec_file=count_file, grna_col="sgRNA", gene_col="gene", count_prior=32., normalization='median', window=800, reference_sample="CTRL")
```
### Run JACKS
JACKS model runs on one gene at a time. To apply JACKS across the measured lines, loop over the desired genes one at a time. JACKS assumes that the measured log2-fold changes $y$ can be obtained as a product of gene essentiality $w$ in a cell line, and the gRNA efficacy $x$: $y_{i,g,l} \sim \mathcal{N}(w_{l,g}x_{i,g}, \sigma^2_{i,l,g})$. 
```{r JACKS, warning=FALSE, message=FALSE}
test_genes = c("KRAS", "RRM2", "ZNF253")
result = infer_jacks(lfc, test_genes)
print(jacks_w_gene(result, "KRAS"), digits=3)
```

### and plot output for one example gene.
The output can be inspected with plot_jacks function, where the raw data, as well as posteriors of $w$ and $x$ are given for all gRNAs and lines:
```{r plot, echo=TRUE, warning=FALSE}
p = plot_jacks(result, "KRAS", do_save=FALSE)
```
 
### Or run JACKS using pre-set reference gRNA efficacy values
JACKS can also be run using a set of precomputed reference values for the gRNAs.
```{r jacks_ref, warning=FALSE, message=FALSE}
result = infer_jacks(lfc, test_genes, reference_library="yusa_v10")
print(jacks_w_gene(result, "KRAS"), digits=3)
